---
title: "Car Price Prediction"
author: "Mohammad Amir"
date: "09/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(caret)
library(ggplot2)
library(gridExtra)
library(dplyr)
```

# Load Data
```{r}
df <- read.csv("Vehicle_Retail_Price_Assignment.csv")
```

# Summary Statistics of Data
```{r}
dim(df) #number of rows and columns in data
head(df) #first 6 rows of data
str(df) #structure of data
summary(df) #statistical summary of data
```

# Check for Missing Data
```{r}
#checking for NAs
colSums(is.na(df)) #there is no NA's in data
```

# Convert Categorical Variables to Factor
```{r}
#convert character variables in factor
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], 
                                       as.factor)
df$symboling <- as.factor(df$symboling)
str(df)
```

# Exploratory Data Analysis

## Exploring the Effect of Car Dimensions on Price 
```{r}
#effect of car dimensions on price
d1 <- ggplot(df, aes(x = wheelbase, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

d2 <- ggplot(df, aes(x = carlength, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

d3 <- ggplot(df, aes(x = carwidth,  y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

d4 <- ggplot(df, aes(x = carheight, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

d5 <- ggplot(df, aes(x = curbweight, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()


grid.arrange(d1, d2, d3, d4, d5, top = "Variation of Price with Car Dimensions")
```

## Engine Technical Specifications on Price
```{r}
#engine technical specifications
et1 <- ggplot(df, aes(x = boreratio, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

et2 <- ggplot(df, aes(x = stroke, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

et3 <- ggplot(df, aes(x = compressionratio, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

et4 <- ggplot(df, aes(x = horsepower, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

et5 <- ggplot(df, aes(x = peakrpm, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

grid.arrange(et1, et2, et3, et4, et5, top = "Price with Technical Engine Specification")
```

## Mean Price Based on Engine Specification
```{r}
c1 <- df %>% group_by(doornumber) %>% summarise(mean_price = mean(price)) %>% {
      ggplot(., aes(x = doornumber, y = mean_price, fill = doornumber)) + 
      geom_bar(stat = "identity") +
      theme_minimal() + theme(legend.position = "none")
}

c2 <- df %>% group_by(carbody) %>% summarise(mean_price = mean(price)) %>% {
  ggplot(., aes(x = carbody, y = mean_price, fill = carbody)) + 
    geom_bar(stat = "identity") +
    theme_minimal() + theme(legend.position = "none")
}

c3 <- df %>% group_by(drivewheel) %>% summarise(mean_price = mean(price)) %>% {
  ggplot(., aes(x = drivewheel, y = mean_price, fill = drivewheel)) + 
    geom_bar(stat = "identity") +
    theme_minimal() + theme(legend.position = "none")
}

c4 <- df %>% group_by(enginelocation) %>% summarise(mean_price = mean(price)) %>% {
  ggplot(., aes(x = enginelocation, y = mean_price, fill = enginelocation)) + 
    geom_bar(stat = "identity") +
    theme_minimal() + theme(legend.position = "none")
}

grid.arrange(c1, c2, c3, c4, top = "Mean Price for Different Configuations")
```

## MPG with Price
```{r}
#mpg with price
mpg1 <- ggplot(df, aes(x = citympg, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

mpg2 <- ggplot(df, aes(x = highwaympg, y = price)) + 
  geom_point() + 
  stat_smooth(method="lm", se=F) + theme_minimal()

grid.arrange(mpg1, mpg2, top = "Price with MPG")
```

## Effect of Symboling on Price
```{r}
#symboling
df %>% group_by(symboling) %>% summarise(mean_price = mean(price)) %>% {
  ggplot(., aes(x = symboling, y = mean_price, fill = symboling)) + 
    geom_bar(stat = "identity") +
    theme_minimal() + theme(legend.position = "none")
}
```

## Boxplots to Show Variation of Price with Different Factors
```{r}
#boxplots
box1 <- df %>% group_by(symboling)  %>% {
        ggplot(., aes(x = symboling, y = price, fill = symboling)) + 
        geom_boxplot() +
        theme_minimal() + theme(legend.position = "none")
}
box1


box2 <- df %>% group_by(fueltype)  %>% {
    ggplot(., aes(x = fueltype, y = price, fill = fueltype)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}

box3 <- df %>% group_by(aspiration)  %>% {
  ggplot(., aes(x = aspiration, y = price, fill = aspiration)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}

box4 <- df %>% group_by(doornumber)  %>% {
  ggplot(., aes(x = doornumber, y = price, fill = doornumber)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}

box5 <- df %>% group_by(carbody)  %>% {
  ggplot(., aes(x = carbody, y = price, fill = carbody)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}

box6 <- df %>% group_by(drivewheel)  %>% {
  ggplot(., aes(x = drivewheel, y = price, fill = drivewheel)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}

box7 <- df %>% group_by(enginelocation)  %>% {
  ggplot(., aes(x = enginelocation, y = price, fill = enginelocation)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}

box8 <- df %>% group_by(enginetype)  %>% {
  ggplot(., aes(x = enginetype, y = price, fill = enginetype)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}

box9 <- df %>% group_by(cylindernumber)  %>% {
  ggplot(., aes(x = cylindernumber, y = price, fill = cylindernumber)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}

box10 <- df %>% group_by(fuelsystem)  %>% {
  ggplot(., aes(x = fuelsystem, y = price, fill = fuelsystem)) + 
    geom_boxplot() +
    theme_minimal() + theme(legend.position = "none")
}




```

```{r}
grid.arrange(box4, box5, box6, box7, top = "Variation of Price for Different Configurations")
```

```{r}
grid.arrange(box2, box3, box8, box9, box10, top = "Variation of Price for Different Engine Specifications")
```

# Correlation Analysis
```{r}
#correlation Analysis
numericData <- df[,sapply(df, is.numeric)] #filter all numeric vars
numericData <- numericData[, -c(1, 15)] #drop the id column and dependent var

library(corrplot)
corMat <- cor(numericData) #correlation matrix
corrplot(corMat, method = "number", type = "lower") #plot of corr matrix
highlyCorrelated <- findCorrelation(corMat, cutoff = 0.7) #find highly correlated
highlyCorCol <- colnames(numericData)[highlyCorrelated]
highlyCorCol
```

# Feature Selection
<p>Feature selection is done using RFE (Recursive Feature Elimination)</p>

## Preprocess the Data for RFE
```{r}
library(caret)
x <- df[, -c(1, 3, 26)] #drop car_ID, CarName and Price (Dependent Var)

#Convert Fators to Numeric for lmFuncs
x$symboling <- as.numeric(x$symboling)
x$fueltype  <- as.numeric(x$fueltype)
x$aspiration<- as.numeric(x$aspiration)
x$doornumber<- as.numeric(x$doornumber)
x$carbody   <- as.numeric(x$carbody)
x$drivewheel<- as.numeric(x$drivewheel)
x$enginelocation<- as.numeric(x$enginelocation)
x$enginetype<- as.numeric(x$enginetype)
x$cylindernumber<- as.numeric(x$cylindernumber)
x$fuelsystem<- as.numeric(x$fuelsystem)

#Normalize Data
normalization <- preProcess(x)
x <- predict(normalization, x)
x <- as.data.frame(x) #Predictors (Independent Vars)
y <- df[, 26] #Dependent Variable
```

## Apply RFE
```{r}
set.seed(5)
lmProfile2 <- rfe(x, y,
                  sizes = c(10:15, 20, 23),
                  rfeControl = rfeControl(functions = lmFuncs,
                                          rerank = TRUE,
                                          number = 200))

lmpImp <- data.frame(varImp(lmProfile2))
lmpImp <- data.frame(variable = rownames(lmpImp), lmpImp)

ggplot(data=lmpImp, 
       aes(x=reorder(variable, 
                     Overall), 
           y=Overall)) +
  geom_bar(stat="identity") + 
  coord_flip()                 #Variable importance plot


rownames(lmpImp) <- NULL
predictors(lmProfile2) #list of selected variables
ggplot(lmProfile2) #plot of RMSE with number of variables in model
ggplot(lmProfile2, metric = "Rsquared") #plot of r-squared
```

# Random Forest

## Selecting Variables from RFE and Train - Test Split
```{r}
selected_vars <- predictors(lmProfile2)
#list of selected variables from RFE

selected_vars <- append(selected_vars, c("price", "car_ID")) #add price to data
df3 <- df[, selected_vars]
library(dplyr)

#train test split
set.seed(125)

train <- df3 %>% dplyr::sample_frac(.75)
test  <- dplyr::anti_join(df3, train, by = 'car_ID')

train_id <- data.frame(car_ID = train$car_ID)
test_id  <- data.frame(car_ID = test$car_ID)

train <- train[,-14]
test  <- test[, -14]

X_train <- train[, -13]
Y_train <- train[, 13]

X_test <- test[, -13]
Y_test <- test[, 13]

```

## Vanilla Random Forest
```{r}
# Train the model 
library(randomForest)

regr <- randomForest(x = X_train, 
                     y = Y_train)

# Make prediction Train
pred_train <- predict(regr, X_train)

result_train <- data.frame(car_ID = train_id, 
                     price = Y_train,
                     predictions = pred_train)
head(result_train)

# Make prediction Test
pred_test <- predict(regr, X_test)

result_test <- data.frame(car_ID = test_id, 
                     price = Y_test,
                     predictions = pred_test)
head(result_test)

#Performance Matrices
library(Metrics)
print(paste0('Test RMSE: ' , rmse(result_test$price,
                                  result_test$predictions) )) #testRMSE
print(paste0('Train RMSE: ' , rmse(result_train$price,
                                  result_train$predictions) )) #RMSE

print(paste0('Test R2: ' ,
             caret::postResample(result_test$predictions , result_test$price)['Rsquared'] ))

print(paste0('Train R2: ' ,
             caret::postResample(result_train$predictions , result_train$price)['Rsquared'] ))

ggplot(result_test, aes(x = predictions, y = price)) + geom_point()
```

## Visualizing Actual and Predicted
```{r}
library(reshape2)

#On Test Data
melt_pred_test <- melt(result_test, id.vars = "car_ID")

ggplot(melt_pred_test, 
       aes(y = value, 
           x = car_ID,
           colour = variable)) +
      geom_point() +
      geom_line() +
  ggtitle("Actual vs Predicted for Test Data")

#On Train Data
melt_pred_train <- melt(result_train, id.vars = "car_ID")

ggplot(melt_pred_train, 
       aes(y = value, 
           x = car_ID,
           colour = variable)) +
      geom_point() +
      geom_line() +
  ggtitle("Actual vs Predicted for Train Data")
```